---
title: '`bcSeq`: Fast Sequence Alignment for High-throughput shRNA and CRISPR Screen'
output: BiocStyle::html_document
author: Jiaxing Lin
date: '2026-06-08'
vignette: >
  %\VignetteIndexEntry{bcSeq}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}

---

``` {r setup1, include=FALSE}
require(knitr)
```

``` {r setup2, include=FALSE}
options(width=80)  # make the printing fit on the page
set.seed(1121)     # make the results repeatable
stdt<-date()
```

# Introduction

Introduction This document provides comprehensive instructions and
examples for using the `bcSeq` package to perform alignment of CRISPR or
shRNA reads to a library of sequences. The alignment is based on a
'Trie' data structure, a tree like data structure for fast searching.
The algorithm is implemented in C++, and ported to R by `Rcpp`.\
Features of this package include:

1.  $O(N\bar{m})$ computational complexity (where $N$ is the number of
    sequence reads in the sample, and $\bar{m}$ is the average read
    length.)

2.  Short sequences alignment that can be applied to barcode matching
    and similar problems.

3.  Allows errors during the alignment. Supports both hamming and edit
    distance matching with constraints.

4.  Alignment qualities are evaluated and ambiguous alignments are
    resolved using Bayes' classifier.

5.  Support for user-defined matching probability models in evaluating
    the alignment quality.

Note: using the 'Trie' data structure only unique barcode sequences in
the library are kept.

# Basic Example

## Simulate a dummy data set

Example Library and Read simulation Generating a library fasta file:

```{r genlib, eval=TRUE}
lFName    <- "./libFile.fasta"
bases     <- c(rep('A', 4), rep('C',4), rep('G',4), rep('T',4))
numOfBars <- 7
Barcodes  <- rep(NA, numOfBars*2)
for (i in 1:numOfBars){
    Barcodes[2*i-1] <- paste0(">barcode_ID: ", i)
    Barcodes[2*i]   <- paste(sample(bases, length(bases)), collapse = '')
}
write(Barcodes, lFName)
```

Generating a read fastq file:

```{r genRead, eval=TRUE}
rFName     <- "./readFile.fastq"
numOfReads <- 8
Reads      <- rep(NA, numOfReads*4)
for (i in 1:numOfReads){
    Reads[4*i-3] <- paste0("@read_ID_",i)
    Reads[4*i-2] <- Barcodes[2*sample(1:numOfBars,1,
        replace=TRUE, prob=seq(1:numOfBars))]
    Reads[4*i-1] <- "+"
    Reads[4*i]   <- paste(rawToChar(as.raw(
        33+sample(20:30, length(bases),replace=TRUE))),
        collapse='')
}
write(Reads, rFName)
```

## Alignment
Alignment by `bcSeq`\_hamming Alignment using default mapping
probability and output

``` {r defaultAlign, eval=TRUE}
library(Matrix)
library(bcSeq)
ReadFile <- "./readFile.fastq"
BarFile  <- "./libFile.fasta"
outFile  <- "./count.csv"
```

``` {r defaultAlign_test, eval=FALSE}
res <- bcSeq_hamming(ReadFile, BarFile, outFile, misMatch = 2,
    tMat = NULL, numThread = 4, count_only = TRUE )
res <- read.csv(outFile, header=FALSE)
```

``` {r sessinfo, echo=FALSE, include=TRUE, results='asis'}
sessionInfo()
```
